Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-01 paper/abstract.tex

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-words data.  Out-of-sample label prediction is the primary goal of this work; however, improved dimensionality reduction is also of interest.  We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).  We find that the additional signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in comparison to supervised LDA approaches that don't utilize information derived from the structure of the label space.   We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks.  We show improved label prediction performance and evidence that the learned topic model also improves as a result of using this signal.

>>>>

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multi-labeled bag-of-word data.  Examples of
such data include pages and their placement in Web directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to structure-agnostic models. Furthermore, we show evidence
that the resulting HSLDA topics are more descriptive of the underlying data
than sLDA topics, which ignore the label hierarchy.

### 2011-06-01 paper/introduction.tex

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured textual data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs~\citep{SNAP}, and patient records and their associated billing codes.
In this work we show how to combine these two sources of information using a
single model that allows one to automatically categorize new text documents,
suggest labels that might be inaccurate, compute improved similarities between
documents for information retrieval purposes, and more. The models and
techniques that we develop in this paper are applicable in other domains as
well, namely, any unstructured representations of data that have been
hierarchically classified (e.g.~image catalogs with bag-of-feature image
representations).

In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.  These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.  It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.   Consider web retail data.  Web retailers often have both a browse-able product hierarchy and free-text descriptions for all products they sell.   The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.  We hypothesize that such hierarchical labels should, at least in theory, provide better supervision than the simpler unstructured labels previously considered.  Results from applying our model to both medical record and web retail data suggests that this is likely the case.  In particular, we observe gains in our primary goal of out-of-sample label prediction that result specifically from leveraging hierarchical supervision.

The remainder of this paper is structured as follows.   Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.

@@

>>>>

The task of multi-label classification, selecting the k-best labels for a given
instance, has been a topic of research for several years. One simplistic way to
carry out the classification is through a series of independent binary
classifiers, but this ignores the many inherent dependencies among the
labels. Thus, much work has been devoted on incorporating the co-occurence
patterns of the labels into the classification task. In this paper, we focus on
multi-label classification, where the labels are organized in a hierarchical
structure. Scenarios of use include, but are not limited to, placing webpages
into manually curated Internet directories~\cite{DMOZ}, categorizing images
according to a taxonomy, tagging product descriptions with catalogue
information~\cite{SNAP}, and assigning diagnosis codes to clinical
records~\cite{Challenge07}.

There are several challenges entailed in incorporating the hierarchical nature
of labels into the classification task. One pertains to the labeling itself: in
the datasets (especially real-world, noisy ones), for a given label, instances
labeled with it contribute positive instance, but it is unclear how to
determine the negative instances. In particular, how to treat the parent labels
of the selected ones?

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. In particular, we
extend supervised latent Dirichlet allocation (sLDA)~\cite{BleiMcAuliffe2008}
to take advantage of hierarchical supervision. We hypothesize that hierarchical
label information provides more information about labeling than considering
labels as a flat list.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Following the trend observed in supervised topic modeling, we
note that the learned topic models are more representative of the underlying
data in both of our datasets~\cite{BleiMcAuliffe2008}.

The remainder of this paper is as follows. Section~\ref{sec:model}
introduces hierarchically supervised LDA (HSLDA), while
Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
Section~\ref{sec:related_work} reviews related work, and
Section~\ref{sec:experiments} shows results from applying HSLDA to health care
and web retail data.

@@

%Our main contribution is to show how to utilize supervision in the form of
%hierarchical and (often) multiple labelings in a similar manner. Consider web
%retail data. Web retailers often have both a browse-able product hierarchy and
%free-text descriptions for all products they sell. The situation of each
%product in a product hierarchy (often multiply situated) constitutes a
%multiple, hierarchical labeling of the free-text product descriptions. We
%hypothesize that such hierarchical labels should, at least in theory, provide
%better supervision than the simpler unstructured labels previously considered.
%Results from applying our model to both medical record and web retail data
%suggests that this is likely the case. In particular, we observe gains in our
%primary goal of out-of-sample label prediction that result specifically from
%leveraging hierarchical supervision.
